A Pile-up of Rapidly Rotating Main-Sequence Stars with a Broad Range of Ages: \\Further Evidence for Weakened Magnetic Braking

### 2021-11-07 src/ms.tex
Updating manuscript

%% The first piece of markup in an AASTeX v6.x document is the \documentclass
%% command. LaTeX will ignore any data that comes before this command. The 
%% documentclass can take an optional argument to modify the output style.
%% The command below calls the preprint style which will produce a tightly 
%% typeset, one-column, single-spaced document.  It is the default and thus
%% does not need to be explicitly stated.
%%
%% using aastex version 6.3
%\documentclass[linenumbers,preprint2]{aastex631}
\documentclass[twocolumn]{aastex631}

@@

%% Reintroduced the \received and \accepted commands from AASTeX v5.2
%\received{March 1, 2021}
%\revised{April 1, 2021}
%\accepted{\today}

@@

%\\or\\Weakened Magnetic Braking Supported by Rotation Rates of Kepler Exoplanet Hosts}

% Ridge
% Edge
% Ledge
% Cornice

@@

%Gaby Contardo
%Lucy Lu
%Joel Zinn
%Adrian Price-Whelan

\begin{abstract}
We combine stellar surface rotation periods determined from NASA's Kepler mission with spectroscopic parameters derived from the California-Kepler Survey and LAMOST--Kepler survey to demonstrate the existence of a long-period pile-up in the effective temperature--rotation period distribution of main-sequence stars with temperatures exceeding 5800~K. Stars in the pile-up are found to have a wide range of ages ($\sim$2--6~Gyr), meaning that, along the ridge, rotation period is strongly predictive of a star's surface temperature but only weakly predictive of its age. In the effective temperature range of 5000--6250~K the long-period pile-up is well-described by a single-parameter model of constant Rossby number, with a critical value of $\mathrm{Ro_{crit}} \sim$\textbf{XYZ}. The long-period pile-up was previously predicted by \citet{vanSaders2019} as a consequence of weakened magnetic braking, in which stellar spin-down stalls once stars reach a critical Rossby number. These results imply that surface rotation periods have limited predictive power for the age determination of F- and early G-type stars older than $\sim$1--2~Gyr. \textbf{We additionally detect a secondary pile-up associated with the short-period edge of the rotation period distribution.}
\end{abstract}

%% Keywords should appear after the \end{abstract} command. 
%% The AAS Journals now uses Unified Astronomy Thesaurus concepts:
%% https://astrothesaurus.org
%% You will be asked to selected these concepts during the submission process
%% but this old "keyword" functionality is maintained in case authors want
%% to include these concepts in their preprints.
\keywords{Stellar rotation (1629) --- Stellar evolution (1599) --- Stellar magnetic fields (1610) --- F dwarf stars(516)}

%% From the front matter, we move on to the body of the paper.
%% Sections are demarcated by \section and \subsection, respectively.
%% Observe the use of the LaTeX \label
%% command after the \subsection to give a symbolic KEY to the
%% subsection for cross-referencing in a \ref command.
%% You can use LaTeX's \ref and \label commands to keep track of
%% cross-references to sections, equations, tables, and figures.
%% That way, if you change the order of any elements, LaTeX will
%% automatically renumber them.
%%
%% We recommend that authors also use the natbib \citep
%% and \citet commands to identify citations.  The citations are
%% tied to the reference list via symbolic KEYs. The KEY corresponds
%% to the KEY in the \bibitem in the reference list below.

@@

%Similar to the apparent stalling of low-mass stars mentioned above,

%If this is the case, a time-dependent (or, more accurately, Rossby-dependent) magnetic field morphology might explain both the anomalously rapid rotation of old field stars and the existence of a detection edge, as the amplitude of variability should also be connected to the field morphology.

%period-detection algorithms are less efficient for stars with (i) smaller amplitude variations, and (ii) quasi-periodic or stochastic variations. The 
%number, size, contrast, and relative orientations of spots on the surface of stars

@@

\section{Sample Selection} \label{sec:sample}
Below, we describe the samples utilized in this work. All stars characterized here were targets of NASA's Kepler mission \citep{Borucki2010} and have published rotation periods derived from the Kepler data.

%We consider two samples in this work. The first sample is derived from the CKS sample, which itself is a subset of Kepler exoplanet candidate host stars that were spectroscopically characterized in a homogeneous manner. The second sample arises from a cross-match of the Kepler field with the LAMOST spectroscopic survey \textbf{(citation)}.

@@

\iffalse
\begin{figure*}
    \centering
    \includegraphics[width=\linewidth]{ridge.pdf}
    \caption{The \teff-\prot plane for the CKS sample. Point colors are scaled to the CKS ages determined from isochrone fitting. The source of \prot is denoted above each panel, where \citet{David2021} is a compilation of vetted periods, rather than a source of original measurements. The black trapezoid indicates the approximate area of the ridge. The grey curves indicate empirical cluster sequences from \citet{Curtis2020}, corresponding to ages of $\sim$2.7, 1, 0.67, and 0.12~Gyr from top to bottom.}
    \label{fig:ridge}
\end{figure*}
\fi

%Points on contours
%Different color for 
%, pixel

@@

We matched 10844 LAMOST targets to 10650 unique Kepler IDs. For the Kepler sources with duplicate cross-matched LAMOST sources we kept the source with a brighter Gaia DR2 $G$ magnitude. We additionally selected sources with 3000~K$< \teff <$~8000~K, 3$<$~\logg~$<$5 dex, and -2~$<$[Fe/H]$<$~2 dex. The final sample contains 10516 unique targets with well-determined \prot and \teff (having a median error of 24~K). There is negligible overlap (only 3 stars) between our LAMOST--Kepler samples and the CKS samples since the \citet{McQuillan2014} did not publish rotation periods for KOIs, which were the targets of the CKS project.

%\textbf{Jason: can you provide any relevant background info and describe the rationale behind the cuts you made to ensure the cross-match was relatively good?}

@@

\subsection{APOGEE--Kepler}
The Apache Point Observatory Galactic Evolution Experiment \citep[APOGEE,][]{Majewski2017} is a large-scale, high-resolution ($R \sim 22500$) stellar spectroscopic survey conducted at $H$-band as part of the Sloan Digital Sky Survey \citep[SDSS-IV][]{Blanton2017}. The spectroscopic analysis pipeline for SDSS DR16 is described in \citet{Jonsson2020}. We used Gaia DR2 source IDs \citep{Gaia2016, Gaia2018} to cross-match Megan Bedell's Gaia--Kepler catalog\footnote{\url{https://gaia-kepler.fun/}} with the APOGEE DR16 catalog \citep{Ahumada2020}. Kepler IDs were then used to cross-match this table with the \citet{McQuillan2014} catalog. While the current overlap between Kepler targets and APOGEE is small compared to the LAMOST catalog, APOGEE DR17 will contain more dwarf stars and provide a better resource for studies such as ours. The focus of this work are overdensities in the \teff--\prot plane, and as these appear to be less prominent when using  APOGEE DR16 temperatures (Figure~\ref{fig:surveys}) we conclude that LAMOST and CKS provide more accurate estimates of \teff and do not analyze the APOGEE sample further.

@@

%\textbf{Measurements of differential rotation on the solar surface have found (SOHO/MDI results and possibly GONG...see e.g. Ulrich 1988)...}

@@

The location of the pile-up approximately corresponds with an empirical hybrid cluster sequence derived by \citet{Curtis2020} from members of the NGC~6819 (age~$\sim$~2.5~Gyr) and Ruprecht~147 (age~$\sim$~2.7~Gyr) open clusters (Figure~\ref{fig:ridge}). As we show in \S\ref{subsec:ages}, stars along the long-period edge have a range of ages. Therefore, the observation that the edge approximately corresponds with the $\sim$2.5--2.7~Gyr cluster sequence implies that stars with \teff$\lesssim 5800~K$ have already piled up onto the edge by or before this timescale. Similarly, the edge is well-separated from the empirical $\sim$1~Gyr cluster sequence based on rotation rates in the NGC~6811 cluster \citep{Curtis2020}, implying that it takes F-type stars $>1$~Gyr to converge onto the long-period edge. These observations are in accordance with predictions from the WMB model which suggest the pile-up forms on a timescale of 2--3~Gyr \citep{vanSaders2019}.

Interestingly, in both the CKS and LAMOST samples, there appears to be a change of slope along the ridge, with an inflection point corresponding closely to the Kraft break at $\approx$6250~K, the point at which convective envelopes become vanishingly thin \citep{Kraft1967}. %\textbf{What to make of this? We can fit a model to the period distribution to measure exactly where the break is and test its significance.}

@@

\iffalse
\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{pctl-curves.pdf}
    \caption{At top, The LAMOST--Kepler sample and the 10th and 90th \prot percentiles, computed as described in \S\ref{subsec:rossby}. At bottom, the same curves are shown in relation to fiducial constant Rossby models.}
    \label{fig:pctl}
\end{figure}
\fi

@@

The constant Rossby model provides a reasonably good description of the long-period edge in the \teff range of $\approx$5000--6250~K, with fractional residuals $\lesssim5\%$ over this range. Above and below this \teff range we see clear and significant divergence from the constant Rossby model, such that the model under-predicts periods of hotter stars and over-predicts periods of cooler stars, because the cooler stars have not had enough time to evolve to that Rossby number \citep[see Figure 6 in][]{vanSaders2019}.

\iffalse
\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{fits.pdf}
    \caption{Constant Rossby model fits to the 90th and 10th \prot percentile curves (top). Residuals from the median models from the MCMC sampling are shown in the lower panel.}
    \label{fig:fits}
\end{figure}
\fi

@@

\iffalse
\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{asteroseismic.pdf}
    \caption{Comparison of CKS (left) and LAMOST--Kepler samples (right) with the \citet{Hall2021} main-sequence asteroseismic sample in the \teff--\prot plane.}
    \label{fig:asteroseismic}
\end{figure}
\fi

@@

Notably, both models fail to reproduce the observed short-period edge.

%\textbf{Some caveats about how the observations and models may not have exactly the same logg distributions? Jen, do you have any guidance here? I think you mentioned being able to get additional parameters for the model populations.}

\iffalse
\begin{figure*}
    \centering
    \includegraphics[width=0.49\linewidth]{std-model-cks.pdf}
    \includegraphics[width=0.49\linewidth]{wmb-model-cks.pdf}
    \includegraphics[width=0.49\linewidth]{std-model-lamost.pdf}
    \includegraphics[width=0.49\linewidth]{wmb-model-lamost.pdf}
    \caption{The \teff-\prot plane for the CKS sample (top panels) and LAMOST sample (bottom panels) in comparison to the standard and WMB models presented in \citet{vanSaders2019}.}
    \label{fig:models}
\end{figure*}
\fi

\iffalse
\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{model-pctl.pdf}
    \caption{Comparison of the 10th and 90th \prot percentiles predicted from the standard (light blue) and WMB (blue) models of \citet{vanSaders2019} with the same percentile ranges from the LAMOST--Kepler observations.}
    \label{fig:model-pctl}
\end{figure}
\fi

@@

% \textbf{Can we measure where the overdensity vanishes by fitting multi-component gaussians to the prot distributions?}

\subsection{Ages of stars on the ridge}
\label{subsec:ages}
The WMB model predicts that hotter stars pile up on the long-period edge at younger ages, producing an age gradient across the edge. In Figure~\ref{fig:ages} we show the \teff-age distributions of stars on the ridge using ages from the CKS and SPOCS \citep{Brewer2018} catalogs. Both catalogs derive spectroscopic parameters from the CKS spectra, but use different pipelines for both the spectroscopic parameters and the isochrone fitting. In both cases, there appears to be an age gradient such that hotter stars are younger on average. However, we note that such a trend is also expected in a sample of main-sequence stars as a natural consequence of the shorter main-sequence lifetimes of hotter, more massive stars. We also find that the dispersion in age is a sensitive function of \teff, with cooler stars on the ridge showing a broader range of ages.

We determined that 90\% of the stars on the ridge have ages between 1.4--5.6~Gyr (using ages from the CKS catalog), or 2.3--5.9~Gyr (using SPOCS ages). In the WMB model, although wind-driven angular momentum losses cease, stars continue to evolve structurally which results in evolution in the moment of inertia and stellar spin, driven by expansion of the stars away from the main-sequence. As a result, an age-gradient is expected along the pile-up as stars of different mass depart the feature at different age. Though we have not measured this gradient, it can be deduced that some stars spend several Gyr occupying the ridge with only modest evolution of their spin rates.

%\textbf{In this study we have allowed for the ridge to have a width, and since stars are not expected to completely stop spinning down in the WMB model, we expect there to be an age-gradient on the ridge along the period axis.}

\iffalse
\begin{figure*}
    \centering
    \includegraphics[width=\linewidth]{ages.pdf}
    \caption{Above, H-R diagram placement of Rossby ridge stars relative to the Sun and the CKS sample (a) and similarly for the LAMOST--Kepler sample (b). Below, the \teff-age plane for CKS stars along the ridge using isochrone ages from the CKS (c) and SPOCS (d) catalogs.}
    \label{fig:ages}
\end{figure*}
\fi

%\section{Discussion} \label{sec:discussion}

@@

\begin{enumerate}
    \item We observe an overdensity in the \teff-\prot plane at the long-period edge for main-sequence stars with \teff~$\gtrsim$5800~K. We hypothesize that this pile-up was previously obfuscated by imprecise \teff estimates. Both the existence of the pile-up and its obscuration by large \teff errors were predicted by \citet{vanSaders2019} as a consequence of weakened magnetic braking for stars with $M\gtrsim1$~\msun.

    \item We find that the long-period pile-up is reasonably well-described by a constant Rossby number, with a critical value of $\mathrm{Ro_{crit}=1.467 \pm 0.007}$, in the \teff range of $\approx$5000-6250~K. A pile-up of stars with a constant Rossby number and a variety of ages is a prediction of the WMB model of \citet{vanSaders2016, vanSaders2019}.

    \item The pile-up is populated by stars with a wide range of isochrone ages ($\sim$1--6~Gyr). Comparison of the feature with empirical rotation sequences from open clusters implies that stars with $M\gtrsim1$~\msun pile-up onto the ridge on a timescale of 1--3~Gyr, compatible with the predictions of \citet{vanSaders2019}. 
    While cooler stars appear to populate the pile-up until an age of $\sim$6~Gyr, the Sun, with an equatorial rotation period of $\approx$25~d and age of 4.567~Gyr, \textbf{appears to have evolved away from the pile-up. Or did it? Not clear}

@@

    % \item We compared our data to the standard and weakened magnetic braking models presented in \citet{vanSaders2019}, \textbf{finding...}

@@

\iffalse
\begin{figure*}
    \centering
    \includegraphics[width=\linewidth]{scatter-color.pdf}
    \caption{LAMOST--Kepler sample in \teff--\prot plane color coded by \logg (left) and [Fe/H] (right). \textbf{This plot shows overlap between the ridge and lower gravity subgiants evolving through it (we think).}}
    \label{fig:my_label}
\end{figure*}
\fi

\iffalse
\begin{figure*}
    \centering
    \includegraphics[width=\linewidth]{tp-comparison.pdf}
    \caption{Comparison of the \teff--\prot distribution for the CKS sample using rotation periods and \teff from the sources indicated by the axes labels.}
    \label{fig:tp-comparison}
\end{figure*}
\fi

\begin{acknowledgments}
We thank Adrian Price-Whelan for providing the APOGEE--Kepler cross-match catalog. It is a pleasure to thank the Astronomical Data Group at the Flatiron Institute and the Stars Group at the American Museum of Natural History for helpful discussions. This work made use of the gaia-kepler.fun crossmatch database created by Megan Bedell. This paper includes data collected by the Kepler mission and obtained from the MAST data archive at the Space Telescope Science Institute (STScI). Funding for the Kepler mission is provided by the NASA Science Mission Directorate. STScI is operated by the Association of Universities for Research in Astronomy, Inc., under NASA contract NAS 5–26555. Guoshoujing Telescope (the Large Sky Area Multi-Object Fiber Spectroscopic Telescope LAMOST) is a National Major Scientific Project built by the Chinese Academy of Sciences. Funding for the project has been provided by the National Development and Reform Commission. LAMOST is operated and managed by the National Astronomical Observatories, Chinese Academy of Sciences. This work has made use of data from the European Space Agency (ESA) mission {\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed by the {\it Gaia} Data Processing and Analysis Consortium (DPAC, \url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding for the DPAC has been provided by national institutions, in particular the institutions participating in the {\it Gaia} Multilateral Agreement. Funding for the Sloan Digital Sky Survey IV has been provided by the Alfred P. Sloan Foundation, the U.S. Department of Energy Office of Science, and the Participating Institutions. SDSS-IV acknowledges support and resources from the Center for High Performance Computing  at the University of Utah. The SDSS website is www.sdss.org. SDSS-IV is managed by the Astrophysical Research Consortium for the Participating Institutions of the SDSS Collaboration including the Brazilian Participation Group, the Carnegie Institution for Science, Carnegie Mellon University, Center for Astrophysics | Harvard \& Smithsonian, the Chilean Participation Group, the French Participation Group, Instituto de Astrof\'isica de Canarias, The Johns Hopkins University, Kavli Institute for the Physics and Mathematics of the Universe (IPMU) / University of Tokyo, the Korean Participation Group, Lawrence Berkeley National Laboratory, Leibniz Institut f\"ur Astrophysik Potsdam (AIP),  Max-Planck-Institut f\"ur Astronomie (MPIA Heidelberg), Max-Planck-Institut f\"ur Astrophysik (MPA Garching), Max-Planck-Institut f\"ur Extraterrestrische Physik (MPE), National Astronomical Observatories of China, New Mexico State University, New York University, University of Notre Dame, Observat\'ario Nacional / MCTI, The Ohio State University, Pennsylvania State University, Shanghai Astronomical Observatory, United Kingdom Participation Group, Universidad Nacional Aut\'onoma de M\'exico, University of Arizona, University of Colorado Boulder, University of Oxford, University of Portsmouth, University of Utah, University of Virginia, University of Washington, University of Wisconsin, Vanderbilt University, and Yale University. This research has made use of NASA's Astrophysics Data System Bibliographic Services.

>>>>

%% The first piece of markup in an AASTeX v6.x document is the \documentclass
%% command. LaTeX will ignore any data that comes before this command. The 
%% documentclass can take an optional argument to modify the output style.
%% The command below calls the preprint style which will produce a tightly 
%% typeset, one-column, single-spaced document.  It is the default and thus
%% does not need to be explicitly stated.
%%
%% using aastex version 6.3
\documentclass[linenumbers,tighten,trackchanges,twocolumn]{aastex631}

@@

%% Reintroduced the \received and \accepted commands from AASTeX v5.2
\received{\today}
%\revised{\today}
%\accepted{\today}

@@

@@

%Lucy Lu
%Joel Zinn
%Adrian Price-Whelan

\begin{abstract}
We combine stellar surface rotation periods determined from NASA's Kepler mission with spectroscopic parameters derived from the California-Kepler Survey and LAMOST--Kepler survey to demonstrate the existence of a long-period pile-up in the effective temperature--rotation period distribution of main-sequence stars with temperatures exceeding 5800~K. Stars in the long-period pile-up are found to have a wide range of ages ($\sim$2--6~Gyr), meaning that, along the ridge, rotation period is strongly predictive of a star's surface temperature but only weakly predictive of its age. In the effective temperature range of 5000--6250~K the long-period pile-up is well-described by a single-parameter model of constant Rossby number, with a critical value of $\mathrm{Ro_{crit}} \sim$\textbf{XYZ}. The long-period pile-up was previously predicted by \citet{vanSaders2019} as a consequence of weakened magnetic braking, in which stellar spin-down stalls once stars reach a critical Rossby number. These results imply that surface rotation periods have limited predictive power for the age determination of F- and early G-type stars older than $\sim$1--2~Gyr. \textbf{We additionally detect a secondary pile-up associated with the short-period edge of the rotation period distribution.}
\end{abstract}

\keywords{Stellar rotation (1629) --- Solar analogs (1941) --- Stellar evolution (1599) --- Stellar magnetic fields (1610) --- Stellar winds(1636)}

@@

@@

\section{Sample Selection} \label{sec:sample}
Below, we describe the samples utilized in this work. All stars characterized here were targets of NASA's Kepler mission \citep{Borucki2010} and have published rotation periods derived from the Kepler data. For each subsample of the Kepler field described below, we combined published rotation periods from a variety of literature sources with spectroscopic parameters provided by large-scale surveys.

@@

\begin{figure*}
    \centering
    \includegraphics[width=\linewidth]{ridge.pdf}
    \caption{The \teff-\prot plane for the CKS sample. Point colors are scaled to the CKS ages determined from isochrone fitting. The source of \prot is denoted above each panel, where \citet{David2021} is a compilation of vetted periods, rather than a source of original measurements. The black trapezoid indicates the approximate area of the ridge. The grey curves indicate empirical cluster sequences from \citet{Curtis2020}, corresponding to ages of $\sim$2.7, 1, 0.67, and 0.12~Gyr from top to bottom.}
    \label{fig:ridge}
\end{figure*}

@@

We matched 10844 LAMOST targets to 10650 unique Kepler IDs. For the Kepler sources with duplicate cross-matched LAMOST sources we kept the source with a brighter Gaia DR2 $G$ magnitude. We additionally selected sources with 3000~K$< \teff <$~8000~K, 3$<$~\logg~$<$5 dex, and -2~$<$[Fe/H]$<$~2 dex. The final sample contains 10516 unique targets with well-determined \prot and \teff (having a median error of 24~K). There is negligible overlap (3 stars) between our LAMOST--Kepler samples and the CKS samples since the \citet{McQuillan2014} did not publish rotation periods for KOIs, which were the targets of the CKS project.

@@

\subsection{APOGEE--Kepler}
The Apache Point Observatory Galactic Evolution Experiment \citep[APOGEE,][]{Majewski2017} is a large-scale, high-resolution ($R \sim 22500$) stellar spectroscopic survey conducted at $H$-band as part of the Sloan Digital Sky Survey \citep[SDSS-IV][]{Blanton2017}. The spectroscopic analysis pipeline for SDSS DR16 is described in \citet{Jonsson2020}. We used Gaia DR2 source IDs \citep{Gaia2016, Gaia2018} to cross-match Megan Bedell's Gaia--Kepler catalog\footnote{\url{https://gaia-kepler.fun/}} with the APOGEE DR16 catalog \citep{Ahumada2020}. Kepler IDs were then used to cross-match this table with the \citet{McQuillan2014} catalog. While the current overlap between Kepler targets and APOGEE is small compared to the LAMOST catalog, APOGEE DR17 will contain more dwarf stars and provide a better resource for studies such as ours. The focus of this work are overdensities in the \teff--\prot plane, and as these appear to be less prominent when using  APOGEE DR16 temperatures (Figure~\ref{fig:surveys}) we conclude that LAMOST and CKS provide more precise estimates of \teff and do not analyze the APOGEE sample further.

@@

@@

The location of the pile-up approximately corresponds with an empirical hybrid cluster sequence derived by \citet{Curtis2020} from members of the NGC~6819 (age~$\sim$~2.5~Gyr) and Ruprecht~147 (age~$\sim$~2.7~Gyr) open clusters (Figure~\ref{fig:ridge}). As we show in \S\ref{subsec:ages}, stars along the pile-up have a range of ages. Therefore, the observation that the edge approximately corresponds with the $\sim$2.5--2.7~Gyr cluster sequence implies that stars with \teff$\lesssim 5800~K$ have already piled up onto the edge by or before this timescale. Similarly, the edge is well-separated from the empirical $\sim$1~Gyr cluster sequence based on rotation rates in the NGC~6811 cluster \citep{Curtis2020}, implying that it takes F-type stars $>1$~Gyr to converge onto the long-period edge. These observations are in accordance with predictions from the WMB model which suggest the pile-up forms on a timescale of 2--3~Gyr \citep{vanSaders2019}.

In both the CKS and LAMOST samples, there is a change of slope along the ridge, with an inflection point corresponding closely to the Kraft break at $\approx$~6250~K, the point at which convective envelopes become vanishingly thin \citep{Kraft1967}. A piecewise linear fit to the ridge in the CKS sample confirmed that the inflection point occurs at \teff~=~6250~$\pm$32~K, where the uncertainty was estimated from Markov chain Monte Carlo (MCMC) sampling. \textbf{Some interpretation.}

@@

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{pctl-curves.pdf}
    \caption{At top, The LAMOST--McQuillan sample and the 10th and 90th \prot percentiles, computed as described in \S\ref{subsec:rossby}. At bottom, the same curves are shown in relation to fiducial constant Rossby models.}
    \label{fig:pctl}
\end{figure}

@@

The constant Rossby model provides a reasonably good description of the long-period edge in the \teff range of $\approx$~5000--6250~K, with fractional residuals $\lesssim5\%$ over this range. Above and below this \teff range we see clear and significant divergence from the constant Rossby model, such that the model under-predicts periods of hotter stars and over-predicts periods of cooler stars, because the cooler stars have not had enough time to evolve to that Rossby number \citep[see Figure 6 in][]{vanSaders2019}.

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{fits.pdf}
    \caption{Constant Rossby model fits to the 90th and 10th \prot percentile curves (top). Residuals from the median models from the MCMC sampling are shown in the lower panel.}
    \label{fig:fits}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{asteroseismic.pdf}
    \caption{Comparison of CKS (left) and LAMOST--Kepler samples (right) with the \citet{Hall2021} main-sequence asteroseismic sample in the \teff--\prot plane. \textbf{Note about teff scales and shifts}.}
    \label{fig:asteroseismic}
\end{figure}

@@

\textbf{Notably, both models fail to reproduce the observed short-period edge.}

\begin{figure*}
    \centering
    \includegraphics[width=0.49\linewidth]{std-model-cks.pdf}
    \includegraphics[width=0.49\linewidth]{wmb-model-cks.pdf}
    \includegraphics[width=0.49\linewidth]{std-model-lamost.pdf}
    \includegraphics[width=0.49\linewidth]{wmb-model-lamost.pdf}
    \caption{The \teff-\prot plane for the CKS sample (top panels) and LAMOST sample (bottom panels) in comparison to the standard and WMB models presented in \citet{vanSaders2019}.}
    \label{fig:models}
\end{figure*}

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{model-pctl.pdf}
    \caption{Discrepancies between the observed and theoretically predicted \teff--\prot distribution of Kepler stars. Comparison of the 10th and 90th \prot percentiles predicted from the standard (light blue) and WMB (blue) models of \citet{vanSaders2019} with the same percentile ranges from the LAMOST--McQuillan sample (black points).}
    \label{fig:model-pctl}
\end{figure}

@@

\subsection{Ages of stars on the ridge}
\label{subsec:ages}
The WMB model predicts that hotter stars pile up on the long-period edge at younger ages, producing an age gradient across the edge. In Figure~\ref{fig:ages} we show the \teff-age distributions of stars on the ridge using ages from the CKS and SPOCS \citep{Brewer2018} catalogs. Both catalogs derive spectroscopic parameters from the CKS spectra, but use different pipelines for both the spectroscopic parameters and the isochrone fitting. In both cases, there appears to be an age gradient such that hotter stars are younger on average. However, we note that such a trend is also expected in a sample of main-sequence stars as a natural consequence of the shorter main-sequence lifetimes of hotter, more massive stars. We also find that the dispersion in age is a sensitive function of \teff, with cooler stars on the ridge showing a broader range of ages. This observation could be due to cooler stars populating the ridge for longer periods of time (relative to hotter stars), the lower precision of isochrone ages for cooler stars, or some combination of the two effects.

We determined that 90\% of the stars on the ridge have ages between 1.4--5.6~Gyr (using ages from the CKS catalog), or 2.3--5.9~Gyr (using SPOCS ages). In the WMB model, although wind-driven angular momentum losses cease, stars continue to evolve structurally which results in evolution in the moment of inertia and stellar spin, driven by expansion of the stars away from the main-sequence. As a result, an age-gradient is expected along the pile-up as stars of different mass depart the pile-up at different ages. Though we have not measured this gradient, the data suggest that some stars spend several Gyr occupying the ridge with only modest evolution of their spin rates.

\begin{figure*}
    \centering
    \includegraphics[width=\linewidth]{ages.pdf}
    \caption{Above, H-R diagram placement of Rossby ridge stars relative to the Sun and the CKS sample (a) and similarly for the LAMOST--Kepler sample (b). Below, the \teff-age plane for CKS stars along the ridge using isochrone ages from the CKS (c) and SPOCS (d) catalogs.}
    \label{fig:ages}
\end{figure*}

@@

\begin{enumerate}
    \item We observe an overdensity at the long-period edge of the \teff-\prot distribution of Kepler main-sequence stars with \teff~$\gtrsim$~5800~K. We hypothesize that this pile-up was previously obfuscated by imprecise \teff estimates. Both the existence of the pile-up and its obscuration by large \teff errors were predicted by \citet{vanSaders2019} as a consequence of weakened magnetic braking for stars with $M\gtrsim1$~\msun.

    \item We find that the long-period pile-up is reasonably well-described by a constant Rossby number, with a critical value of $\mathrm{Ro_{crit}=1.467 \pm 0.007}$, in the \teff range of $\approx$~5000-6250~K. A pile-up of stars with a constant Rossby number and a variety of ages is a prediction of the WMB model of \citet{vanSaders2016, vanSaders2019}.

    \item The pile-up is populated by stars with a wide range of isochrone ages ($\sim$1--6~Gyr). Comparison of the feature with empirical rotation sequences from open clusters implies that stars with $M\gtrsim1$~\msun pile-up onto the ridge on a timescale of 1--3~Gyr, compatible with the predictions of \citet{vanSaders2019}. 
    While cooler stars appear to populate the pile-up until an age of $\sim$6~Gyr, the Sun, with an equatorial rotation period of $\approx$25~d and age of 4.567~Gyr, may have evolved away from the pile-up. \textbf{However, this statement depends on the magnitude of any shift between the \teff scales of the Sun and the spectroscopic surveys used in this work, as well as any shift between the age scales of the solar system and the theoretical isochrones used to age-date the CKS sample. For example, stars which populate the ridge with isochrone ages older than the Sun may in fact have ages younger than or equal to the Sun's if systematic offsets exist.}

    %Ruth: include some discussion of what this says about the Sun. Not just because it implies the Sun's dynamo might be different from McQuillan stars because it's been through WMB, but also how did the Sun get there? Its Rossby number is higher than the stars on the ridge... Can expansion alone explain the ~6 days of rotation period lengthening it has gone through since it passed through the ridge around 2.5 Gyr ago?

@@

@@

\begin{figure*}
    \centering
    \includegraphics[width=\linewidth]{scatter-color.pdf}
    \caption{LAMOST--Kepler sample in \teff--\prot plane color coded by \logg (left) and [Fe/H] (right). \textbf{This plot shows overlap between the ridge and lower gravity subgiants evolving through it (we think).}}
    \label{fig:my_label}
\end{figure*}

\begin{figure*}
    \centering
    \includegraphics[width=\linewidth]{comparison.pdf}
    \caption{Comparison of the \teff--\prot distribution for the CKS sample using rotation periods and \teff from the sources indicated by the axes labels.}
    \label{fig:tp-comparison}
\end{figure*}

\begin{acknowledgments}
We thank Adrian Price-Whelan for providing the APOGEE--Kepler cross-match catalog and Rodrigo Luger for assistance with the \texttt{showyourwork} package. It is a pleasure to thank the Astronomical Data Group at the Flatiron Institute and the Stars Group at the American Museum of Natural History for helpful discussions. This work made use of the gaia-kepler.fun crossmatch database created by Megan Bedell. This paper includes data collected by the Kepler mission and obtained from the MAST data archive at the Space Telescope Science Institute (STScI). Funding for the Kepler mission is provided by the NASA Science Mission Directorate. STScI is operated by the Association of Universities for Research in Astronomy, Inc., under NASA contract NAS 5–26555. Guoshoujing Telescope (the Large Sky Area Multi-Object Fiber Spectroscopic Telescope LAMOST) is a National Major Scientific Project built by the Chinese Academy of Sciences. Funding for the project has been provided by the National Development and Reform Commission. LAMOST is operated and managed by the National Astronomical Observatories, Chinese Academy of Sciences. This work has made use of data from the European Space Agency (ESA) mission {\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed by the {\it Gaia} Data Processing and Analysis Consortium (DPAC, \url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding for the DPAC has been provided by national institutions, in particular the institutions participating in the {\it Gaia} Multilateral Agreement. Funding for the Sloan Digital Sky Survey IV has been provided by the Alfred P. Sloan Foundation, the U.S. Department of Energy Office of Science, and the Participating Institutions. SDSS-IV acknowledges support and resources from the Center for High Performance Computing  at the University of Utah. The SDSS website is www.sdss.org. SDSS-IV is managed by the Astrophysical Research Consortium for the Participating Institutions of the SDSS Collaboration including the Brazilian Participation Group, the Carnegie Institution for Science, Carnegie Mellon University, Center for Astrophysics | Harvard \& Smithsonian, the Chilean Participation Group, the French Participation Group, Instituto de Astrof\'isica de Canarias, The Johns Hopkins University, Kavli Institute for the Physics and Mathematics of the Universe (IPMU) / University of Tokyo, the Korean Participation Group, Lawrence Berkeley National Laboratory, Leibniz Institut f\"ur Astrophysik Potsdam (AIP),  Max-Planck-Institut f\"ur Astronomie (MPIA Heidelberg), Max-Planck-Institut f\"ur Astrophysik (MPA Garching), Max-Planck-Institut f\"ur Extraterrestrische Physik (MPE), National Astronomical Observatories of China, New Mexico State University, New York University, University of Notre Dame, Observat\'ario Nacional / MCTI, The Ohio State University, Pennsylvania State University, Shanghai Astronomical Observatory, United Kingdom Participation Group, Universidad Nacional Aut\'onoma de M\'exico, University of Arizona, University of Colorado Boulder, University of Oxford, University of Portsmouth, University of Utah, University of Virginia, University of Washington, University of Wisconsin, Vanderbilt University, and Yale University. This research has made use of NASA's Astrophysics Data System Bibliographic Services.
